---
title: "Practical Machine Learning: Prediction Assignment Writeup"
author: "Hari"
date: "12 July 2018"
output: html_document
---
## Introduction
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. 
In this project, goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict whether they did exercise correctly

## Load required Libraries
```{r loadlib, include=FALSE}
library(caret)
library(randomForest)
library(doParallel)
```

## Download Input Files 
```{r setup, include=FALSE}
setwd("C:/Users/Hari/Documents/rwork/pracMacLearn" )

trainfl <- "pml_training.csv"
testfl <- "pml_testing.csv"
trainurl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testurl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
if (!file.exists(trainfl)) {
    download.file(trainurl, destfile = trainfl)
}


if (!file.exists(testfl)) {
    download.file(testurl, destfile = testfl)
}


```

## Read the data
```{r Readdata}
training <- read.csv(trainfl, na.strings = c("NA","#DIV/0!",""))
dim(training)

testing <- read.csv(testfl, na.strings = c("NA","#DIV/0!",""))
dim(testing)

```

## Clean data -  
```{r Clean}
#Remove columns with Near Zero Values
subtrain <- training[, names(training)[!(nzv(training, saveMetrics = T)[, 4])]]
# Remove columns with NA or is empty
subtrain <-
    subtrain[, names(subtrain)[sapply(subtrain, function (x)
        ! (any(is.na(x) | x == "")))]]

```

## Including Plots
 

```{r pressure, echo=FALSE}
subtrain <- subtrain[,-1]
subtrain <- subtrain[, c(1:3, 5:58)]

# Divide the training data into a training set and a validation set
intrain <- createDataPartition(subtrain$classe, p = 0.6, list = FALSE)
subtraining <- subtrain[intrain,]
subvalidation <- subtrain[-intrain,]
# Check if model file exists
model <- "modelFit.RData"
if (!file.exists(model)) {

    # If not, set up the parallel clusters.  
    require(parallel)
    require(doParallel)
    cl <- makeCluster(detectCores() - 1)
    registerDoParallel(cl)
    
    fit <- train(subtraining$classe ~ ., method = "rf", data = subtraining)
    save(fit, file = "modelFit.RData")
    
    stopCluster(cl)
} else {
    # Good model exists from previous run, load it and use it.  
    load(file = "modelFit.RData", verbose = TRUE)
}

predtrain <- predict(fit, subtraining)

confusionMatrix(predtrain, subtraining$classe)

predvalidation <- predict(fit, subvalidation)
confusionMatrix(predvalidation, subvalidation$classe)

varImp(fit)
fit$finalModel
predtest <- predict(fit, testing)
predtest

pml_write_files = function(x){
    n = length(x)
    for(i in 1:n){
        filename = paste0("problem_id_",i,".txt")
        write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
    }
}

pml_write_files(predtest)
```
 